Advances in Artificial Intelligence for Data Visualization: Automate Reading of Diagnostic Plots with Compute Vision Models

### 2023-04-11 Milestone_2/report/report.Rmd
modify the desc for CV
completely rewrite current outcome

The workload of individuals could be alleviated by automating repetitive tasks and providing standardized results in a controlled environment. The evaluation of visual tests on a large scale is not feasible without the aid of machines and technology.

@@

- cut this p down, we don't need that many details

In the initial phase of this research, a pilot investigation was undertaken to explore the viability of implementing visual inference in regression diagnostics. The pilot study involved 64 participants and focused on utilizing a lineup consisting of 20 plots, wherein a residual plot was embedded along with 19 null plots, simulated using the residual rotation technique. The study considered two primary forms of residual departures, namely non-linearity and heteroskedasticity, within a multiple linear regression context. To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete uniform, were also incorporated.

The outcomes of this pilot study demonstrated that the power of visual tests in regression diagnostics increases proportionately with the number of lineup evaluations. Nevertheless, it was observed that conventional tests, such as the F-test for non-linearity and the Breusch-Pagan test for heteroskedasticity, exhibited significantly greater power than the visual tests. While the pilot study successfully provided a comparative analysis of the power of visual and conventional tests, the impact of various factors on the power of visual tests could not be fully analyzed due to inadequate data.

We acknowledge the limitations of this pilot study, particularly in the design of the simulated model, which made it challenging to explicate the involved factors. Furthermore, the methodology utilized for effect size calculation, p-value calculation, and power estimation was incipient. Hence, a follow-up study is necessary to address these limitations and obtain a more comprehensive understanding of the potential of visual inference in regression diagnostics.

In the subsequent year, a similar study was conducted, utilizing simpler simulated models to elicit more distinct visual patterns. Specifically, the non-linearity model was formulated as

\begin{align} \label{eq:nonlinearity-model}
\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1), \\
\boldsymbol{z} &= g(\boldsymbol{z}_{raw}, 1), \\
\boldsymbol{z}_{raw} &= He_j(g(\boldsymbol{x}, 2)),
\end{align}

\noindent where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vectors of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $[-k, k]^n$ defined as

\begin{equation} \label{eq:scaling-function}
g(\boldsymbol{x}, k) = \frac{\boldsymbol{x} - min(\boldsymbol{x})}{max(\boldsymbol{x}) - min(\boldsymbol{x})}(2k - k), \quad \text{for} \quad k > 0. 
\end{equation}

The heteroskedasticity model was formulated as

\begin{align} \label{eq:heter-model}
\boldsymbol{y} &= 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} &= g(\boldsymbol{x}_{raw}, 1),\\
\boldsymbol{\varepsilon} &\sim N(\boldsymbol{0}, 1 + (2 - |a|)(\boldsymbol{x} - a)^2b \boldsymbol{I}). 
\end{align}

The null regression model used to fit the realizations generated by the above models is formulated as

\begin{equation} \label{eq:null-model}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

\noindent where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.

Since $z = O(x^j)$, for $j > 1$, $z$ is a higher order term leaves out by the null regression, which will violate the linearity model assumption. Similarly, for $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the predictor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity.

- draft below

The effect size was derived using the Kullback-Leibler divergence method, formulated as Equation 3. To account for the interdependence of lineups presented to different participants, p-value calculation, as suggested by VanderPlas et al. (2021), was adopted. Furthermore, power estimation was simplified through a logistic regression model that lacked an intercept but included a fixed offset to ensure that the power equated to 5% when the effect size was zero. A detailed description of the experimental design, effect size derivation, p-value calculation, and power estimation can be found in the attached paper.

3. summarise the work in the first year
- an experiment to compare visual test with conventional test 
- a prototype model for lineup evaluation
- points out the limitation of the work
4. summarise the work in the second year
- refine and reconduct the experiment
5. main findings to date
- findings from the first year
- findings from the second year

>>>>

@@

In order to examine the potential applicability of integrating visual inference techniques into regression diagnostics, a pilot study was conducted in the first year with the participation of 64 individuals, which was followed by a formal study involving 443 participants in the subsequent year. The participants were presented with lineups consisting of 20 plots, where a residual plot was embedded along with 19 null plots, drawn with data simulated using the residual rotation technique. The study considered two primary forms of residual departures in multiple linear regression, namely non-linearity and heteroskedasticity. To enrich the visual features, different fitted value distributions, including normal, uniform, lognormal, and discrete, were also incorporated.

The study revealed that conventional residual-based statistical tests are more sensitive to weak residual departures from model assumptions compared to visual tests evaluated by humans, leading to excessive rejections even when downstream analysis and outcomes would not be significantly impacted by the small departures from a good fit. Specifically, conventional tests conclude that the model fit has issues nearly twice as frequently as humans would, and they frequently reject departures in the form of non-linearity and heteroskedasticity that are not visible to humans. These findings emphasize the crucial role of graphical diagnostics and support the integration of visual inference into regression diagnostics. Moreover, they provide a compelling rationale for the development of an automated visual inference system to evaluate lineups of residual plots. The comprehensive details of the research methodology, along with other intriguing findings can be found in the attached paper.
